---
title: "Student Performance Analysis"
author: "Sanjay Chhetri"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Introduction

The goal of this project is to analyze how students' reading and writing scores relate to their math scores and to build a simple predictive model based on these relationships.

# Data Loading

```{r}
# Load the dataset
student_data <- read_csv("StudentsPerformance.csv")

# View the first few rows
head(student_data)
```

# Data Exploration

```{r}
# Structure of the dataset
glimpse(student_data)

# Summary statistics
summary(student_data)

# Dataset dimensions
dim(student_data)
```

# Visualizations

## Distribution of Math Scores

```{r}
ggplot(student_data, aes(x = `math score`)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Distribution of Math Scores", x = "Math Score", y = "Count")
```

## Reading Score vs Math Score

```{r}
ggplot(student_data, aes(x = `reading score`, y = `math score`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Reading Score vs Math Score", x = "Reading Score", y = "Math Score")
```

## Writing Score vs Math Score

```{r}
ggplot(student_data, aes(x = `writing score`, y = `math score`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Writing Score vs Math Score", x = "Writing Score", y = "Math Score")
```

# Modeling

## Linear Regression Model

```{r}
# Build the linear model
model <- lm(`math score` ~ `reading score` + `writing score`, data = student_data)

# Model summary
summary(model)
```

# Results and Interpretation

- Both reading and writing scores are statistically significant predictors of math score.
- R-squared value is approximately 0.674, meaning about 67.4% of the variability in math scores can be explained by reading and writing scores.

# Conclusions

- Students who perform better in reading and writing are likely to perform better in math.
- Reading and writing scores are strong predictors for math performance.

# Future Work

- Incorporate more features like parental education, lunch status, and test preparation.
- Test model improvements using different regression techniques.
- Explore classification approaches (e.g., pass/fail prediction based on scores).
